\pkg{sdmTMB}: An \proglang{R} Package for Fast, Flexible, and User-Friendly Generalized Linear Mixed Effects Models with Spatial and Spatiotemporal Random Fields

### 2023-09-15 doc/paper-jss/sdmTMB-paper.Rnw
Edits

\begin{landscape}
\begin{figure}[p]
\centering
\includegraphics[width=9in]{../../figs/diagram}
\caption[]{
Components of an \pkg{sdmTMB} model with illustrations, descriptions, examples,
notation, and example code. An \pkg{sdmTMB} model can be built from any
combination of the process components (first six rows) plus an observation
component (last row). The examples are from an SDM context, but the model can be
fit to any spatially referenced point data. Notation: We refer to design
matrices as $\bm{X}$. The indexes $\bm{s}$, $t$, and $g$ index spatial
coordinates, time, and group, respectively. The $\sigma$ and $\bm{\Sigma}$
symbols represent standard deviations and covariance matrices, respectively. All
other symbols refer to the  described model components (e.g., $\bm{\beta}$ and
$\bm{\omega}$ refer to a vector of main effects and spatial random field
deviations, respectively). See the \pkg{sdmTMB}
\href{https://pbs-assess.github.io/sdmTMB/articles/}{model description vignette}
for a full description of the model. Note that \code{s()} denotes a smoother as
in \pkg{mgcv} \citep{wood2017a}, \code{breakpt()} denotes a breakpoint
'hockey-stick' shape \citep[e.g.][]{barrowman2000}, \code{(1|g)} denotes
a random intercept by group \code{g}, and \code{\textasciitilde 0} is used in
an \proglang{R} formula to omit an intercept.}\label{fig:diagram}
\end{figure}
\end{landscape}

@@

We find that the Poisson-link delta-lognormal model is favoured with AIC.

@@

We conducted 50 simulation iterations per model and mesh resolution and show
results for the mean, lower, and upper values (Figure~\ref{fig:timing}). The
models were fit on a 2023 16 inch M2 MacBook Pro with an Apple M2 Pro 10-core
CPU and 32 GB of RAM in R 4.2.2 and the default Matrix library. The following
package versions were used:

<<pkg-versions, echo=TRUE>>=
packageVersion("INLA")
packageVersion("inlabru")
packageVersion("mgcv")
packageVersion("spaMM")
packageVersion("sdmTMB")
@

>>>>

% \begin{landscape}
\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{../../figs/diagram}
\caption[]{
Components of an \pkg{sdmTMB} model with illustrations, descriptions, examples,
notation, and example code. An \pkg{sdmTMB} model can be built from any
combination of the process components (first six rows) plus an observation
component (last row). The examples are from an SDM context, but the model can be
fit to any spatially referenced point data. Notation: We refer to design
matrices as $\bm{X}$. The indexes $\bm{s}$, $t$, and $g$ index spatial
coordinates, time, and group, respectively. The $\sigma$ and $\bm{\Sigma}$
symbols represent standard deviations and covariance matrices, respectively. All
other symbols refer to the  described model components (e.g., $\bm{\beta}$ and
$\bm{\omega}$ refer to a vector of main effects and spatial random field
deviations, respectively). See the \pkg{sdmTMB}
\href{https://pbs-assess.github.io/sdmTMB/articles/}{model description vignette}
for a full description of the model. Note that \code{s()} denotes a smoother as
in \pkg{mgcv} \citep{wood2017a}, \code{breakpt()} denotes a breakpoint
'hockey-stick' shape \citep[e.g.][]{barrowman2000}, \code{(1|g)} denotes
a random intercept by group \code{g}, and \code{\textasciitilde 0} is used in
an \proglang{R} formula to omit an intercept.}\label{fig:diagram}
\end{figure}
% \end{landscape}

@@

We find that the Poisson-link delta-lognormal model is favoured by AIC.

@@

We conducted 50 simulation iterations per model and mesh resolution and show
results for the mean, lower, and upper values (Figure~\ref{fig:timing}). The
models were fit on a 2023 16 inch M2 MacBook Pro with an Apple M2 Pro 10-core
CPU and 32 GB of RAM in R 4.2.2 and the default Matrix library.

% The following package versions were used:

<<pkg-versions, echo=FALSE, eval=FALSE>>=
packageVersion("INLA")
packageVersion("inlabru")
packageVersion("mgcv")
packageVersion("spaMM")
packageVersion("sdmTMB")
@
